---
title: "Non-Parametric Poisson Regression"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: cerulean
---

This work is going to use $scam$ to achieve non-parametric regression for modeling arterial input function data for PET imaging.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scam)
library(ggplot2)
```

## Readin data
```{r}
filenames = list.files(path = "./data", pattern="*.csv") ##name of files
data = tibble(
  patient = substr(filenames,1,5), #the first 5 character of filenames
  count = map(filenames,~read.csv(paste0("./data/", .)) %>% select(-X))
)
```
## Find tmax and seperate data

$tmax$ is the time point when AIF reaches a maximum. We only use data after $tmax$ in this study.
```{r}
findtmax <- function (data){
  return(data %>% mutate(tmax = time[which(aif == max(aif))]))
}

slice_exp <- function(conc){
  conc = conc %>% 
    filter(time>=time[which(aif == max(aif))]) %>%
    mutate(t_G = t_G-time[which(aif == max(aif))],
           time = time-time[which(aif == max(aif))])
  return(conc)
}

slice_asc <- function(conc){
  conc = conc %>% 
    filter(time<=time[which(aif == max(aif))])
  return(conc)
}

data = data %>% 
  group_by(patient) %>% 
  mutate(count = map(count,~findtmax(.x)), # the tmax is the one with max aif
         count_asc = map(count, ~slice_asc(.x)), # data before tmax
         count_dsc = map(count, ~slice_exp(.x))) # data after tmax and time=time-tmax; t_G=t_G-tmax

```


```{r, include = FALSE}
#c = data$count_dsc[1] %>% as.data.frame()
#plot(c$time, c$count)
```

## Non-parametric regression

Use $scam$ package and $scam$ function to fit model. In this regression, we set $delta$ (time in the gamma tube), $vol$ (volum), $disp$ (dispersion correction), decay correction, calibration, metabolite correction, and blood-to-plasma correction as offsets.
```{r}
non_regress = function(data = data, calibration = 0.003, disp = 1 ){
fit_res = scam(count ~ s(time,k =15, bs="mpd"),
               offset = log(delta)+log(vol)+log(disp)+(-log(2)/20.364*t_G)+(-log(calibration)*rep(1,length(time)))+(-log(parentFraction))+log(bpr)
              ,family = poisson(link = "log"),data = data)

return(fit_res)
}

fit_data = data %>% 
  group_by(patient) %>% 
  mutate(
    dsc_mod = map(count_dsc,~non_regress(.x)),
    dsc_pred =map(dsc_mod,~.x$linear.predictors),
    dsc_res = map(dsc_mod,~.x$residuals)
    )
```

```{r, include = FALSE}
#reg = c %>% non_regress()
#par(mfrow=c(1,2))
#plot(reg,shade=TRUE)
```

## Plots

### some notes for the second plot
- Grey shade: produces shaded regions as confidence bands for smooths.
- Black bar at the bottom: They are data points between 0 and 10 and they are to close so that form a bar.

### some notes for the fourth plot
- The red line is the median of the residuals.

```{r}
for (i in 1:nrow(fit_data)){
   patient = fit_data$patient[i] 
  # plots of line
  data_line = fit_data$count_dsc[i] %>% as.data.frame()
  plot_line = data_line %>% non_regress()
  
  plot(data_line$time, log(data_line$aif), 
       xlab = "time(min)",
       ylab = "log(AIF)",
       main = paste0("Raw data for patient:", patient))
  
  plot(plot_line,shade=TRUE,
       xlab = "time(min)",
       ylab = "log(count/offset)",
       main = "The Component smooth functions on the scale of the linear predictor")
  
 
  #ticklabels = c(1,2,3,4,5,6,7,8,9,10,20,60,80)
  #axis(1, at =ticklabels, lwd = ticklabels)
  
   # plots of prediction
  plot_pred = fit_data$dsc_pred[i] %>% as.data.frame()
  names(plot_pred)[1] = "pred"
  plot(data_line$time, plot_pred$pred, 
       xlab = "time(min)",
       ylab = "log(count)",
       main = paste0("Predicted data for patient:",patient))

  # plots of residuals
  res = fit_data$dsc_res[i]%>% as.data.frame()
  names(res)[1] = "residual"
  boxplot(res$residual,
       main = paste0("Residule for patient:",patient))
  abline(h=median(res$residual),col = "red")
  text(1 - 0.4, median(res$residual), 
       labels = formatC(median(res$residual), format = "f", 
                        digits = 3),
       pos = 3, cex = 0.9, col = "red")
}

```

